Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-01-25 development.tex
added two new sections.
adding a bunch of new sections and subsections.
Couple paragraphs on development.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

\begin{enumerate}
\item near-future development plans
\item far-future development plans
\item clear areas of improvement
\end{enumerate}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code development methodology}

Over time, Enzo's development has followed a trajectory toward openness.
Started as the graduate research project of Greg Bryan at the University of
Illinois, it has been subsequently been stewarded by the Laboratory for
Computational Astrophysics (LCA) at the University of California at San Diego,
and has transitioned to a distributed, completely open, and community-driven
project.  Initially, Enzo was versioned using a series of ``snapshots'' of the
code base, usually hand-created by the developing individuals.  These were
distributed to collaborators and colleagues, but the central ``trunk'' of
development was updated primarily by a single person: while patches and
technology were accepted from external developers, the relatively small number
of individuals using the code resulted in a strong centralization of
development.

As the stewardship of the code passed to the LCA, the code was released first
to ``friendly users'' and then as an open source release.  However, while the
code was made available with documentation, technology developed in the broader
community of users was typically not re-integrated.  This led to a wide
dispersal of development, largely independent, by individuals who downloaded
and used the version of the code developed by the LCA.

Following the first public, open source release of Enzo, the code was migrated
to using the Subversion version control system.  This is a centralized version
control system, and the ``stable'' Enzo source code was made globally readable
following the Enzo 1.5. release.  Access to the primary development tree
required a password and login for each user, and providing upstream changes
either required this password and the granting of write access or a sequence of
patches and manually-created diffs (much like the original development system.)
The technical friction of manually contributing patches and modifications,
combined with Subversion's difficulty with tracking merges, resulted in further
fragmentation of the code base.

The main development streams of Enzo were eventually merged using the
distributed version control system (DVCS) Mercurial
(\url{http://mercurial.selenic.com/}) into a branch of the code known as
\texttt{week-of-code}, so named after the in-person development sprint at which
it was created.  The fundamental, and transformative, distinction between the
previous \textit{centralized} version control system and mercurial's
\textit{distributed} version control system is the elimination of gatekeepers.
While there still exists a canonical, central location that stable and
development versions of Enzo can be obtained, changesets and versions can be
exchanged between peers without the intervention of designated gatekeepers.
This has the direct effect of enabling local development to be versioned and
provenance ensured, while still retaining the ability to benefit from
``upstream'' development.  An important, even crucial, side effect is that the
technology used for local versioning provides mechanisms for submitting
locally-developed modifications to the community source location.  Mercurial
internally represents all changes as nodes in a directed, acyclic graph (DAG),
which results in the natural ability to more consistently and easily manage
merging development streams.

Currently, Enzo is developed using the hosted source control platform Bitbucket
(\url{http://bitbucket.org/}) at \url{http://bitbucket.org/enzo/}.  There are
two mailing lists, one for usage-focused questions and discussion, and another
for development discussion.  Both of these lists are open and publicly
archived.  Bitbucket provides mechanisms for inspecting source, hosting
branches and forks of the primary source, as well as code review.  All proposed
source code changes for Enzo are subjected to a peer review process, where
experienced developers read, inspect, test and provide feedback on source code
changes.  All developers, including long-time Enzo contributors and developers,
are subject to this process for including code in the primary Enzo repository.
By using a remote, hosted system, Enzo is now \textit{completely} open to
contributions from the community.  Individuals, who may or may not consider
themselves Enzo developers, are free to ``fork'' the Enzo code base, develop
changes (signed with their own name), and submit them for review and inclusion.
In contrast to the centralized, gatekeeper-focused technology used previously,
this enables anyone to contribute changes to be evaluated for inclusion in the
Enzo codebase.

While peer review is able to catch many bugs and problems with source
code changes, Enzo is also subject to ``answer'' testing.  We have created a
set of parameter files and problem types that exercise the underlying machinery
of Enzo.  These ``test problems'' have affiliated ``tests,'' which consist of
scripts that use \texttt{yt} \citep{2011ApJS..192....9T} to produce results
such as mass distribution, projections, profiles and so on.  The testing
infrastructure then evaluates whether the variation in the new results compared
to a ``gold standard'' of results has exceeded an acceptable threshold,
typically set to roughly single precision.  Optionally, for those test problems
that are deemed unsafe to change to any precision, the tests also produce
hashes of the outputs; these hashes will only remain unchanged in the event of
bitwise identicality between results.  The results of the gold standard are
versioned and stored in Amazon S3, enabling remote testing to proceed.  While
the testing process -- building, running, analyzing and comparing -- is not yet
automated against incoming pull requests, we hope to deploy that functionality
in the future.  The primary challenge is that of compute time; the tests are
organized into multiple categories, including by the expected run time, but
the full suite of tests can take several days to run.
